---
title: "Assignment 01 - Data Viz and Wrangling"
output: github_document
always_allow_html: true
---

```{r setup, message=FALSE}
library(data.table)
library(leaflet)
library(tidyverse)
library(ggplot2)
library(lubridate)
```



1.

- Download process was done without R and thus is not shown here.

```{r}
indiv <- data.table(read.csv("chs_individual.csv"))
region <- data.table(read.csv("chs_regional.csv"))

merged <- merge(x = indiv, y = region, by = "townname",  all.x = TRUE, all.y = FALSE)

nrow(indiv)
nrow(merged)

merged[, bmi_avg := mean(bmi, na.rm = TRUE), by = .(male, hispanic)]
merged[, fev_avg := mean(fev, na.rm = TRUE), by = .(male, hispanic)]

merged[, bmi := ifelse(is.na(bmi), bmi_avg, bmi)]
merged[, fev := ifelse(is.na(fev), fev_avg, fev)]
```

- Missing values in asthma cannot be dealt with using the provided method because asthma is a binary variable.



2.

```{r, message=FALSE}
merged[bmi <= 14, obesity_level := "underweight"]
merged[bmi <= 22 & bmi > 14, obesity_level := "normal"]
merged[bmi <= 24 & bmi > 22, obesity_level := "overweight"]
merged[bmi > 24, obesity_level := "obese"]

merged %>%
  group_by(obesity_level) %>%
    summarise(min = min(bmi), max = max(bmi), n())
```



3.

```{r}
merged[smoke == 1 & gasstove ==1, smoke_gas_exposure := "both"]
merged[smoke == 1 & gasstove ==0, smoke_gas_exposure := "smoke"]
merged[smoke == 0 & gasstove ==1, smoke_gas_exposure := "gas"]
merged[smoke == 0 & gasstove ==0, smoke_gas_exposure := "neither"]
```



4.

```{r, message=FALSE}
merged %>%
  group_by(townname) %>%
    summarise(avg_fev = mean(fev), sd_fev = sd(fev), prop_as = mean(asthma, na.rm = TRUE), sd_as = sd(asthma, na.rm = TRUE))

merged %>%
  group_by(male) %>%
    summarise(avg_fev = mean(fev), sd_fev = sd(fev), prop_as = mean(asthma, na.rm = TRUE), sd_as = sd(asthma, na.rm = TRUE))

merged %>%
  group_by(obesity_level) %>%
    summarise(avg_fev = mean(fev), sd_fev = sd(fev), prop_as = mean(asthma, na.rm = TRUE), sd_as = sd(asthma, na.rm = TRUE))

merged %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
    group_by(smoke_gas_exposure) %>%
      summarise(avg_fev = mean(fev), sd_fev = sd(fev), prop_as = mean(asthma, na.rm = TRUE), sd_as = sd(asthma, na.rm = TRUE))
```




Looking at the data (EDA)

```{r}
merged %>%
  ggplot(mapping = aes(x=bmi, y=fev, color=townname))+
  geom_point()+
  stat_smooth(method=lm)+
  facet_wrap(~townname)+
  labs(title = "BMI v. FEV by town with regression lines")
```

- BMI and FEV have positive correlation across all towns.

```{r}
merged %>%
  ggplot()+
  geom_histogram(mapping = aes(x=fev, fill=obesity_level))+
  scale_fill_brewer(palette = "Set3")+
  labs(title = "FEV by BMI category", x="FEV", y="n")

merged %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
    ggplot()+
    geom_histogram(mapping = aes(x=fev, fill=smoke_gas_exposure))+
    scale_fill_brewer(palette = "Set3")+
    labs(title = "FEV by smoke/gas exposure", x="FEV", y="n")

```

- Within each category of obesity level or smoke/gas exposure, the distribution of FEV is nearly normal. Most people have normal BMI and exposure to gas.

```{r}
merged %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
  ggplot()+
  geom_bar(mapping = aes(x=obesity_level, fill=smoke_gas_exposure))+
  scale_fill_brewer(palette = "Set3")+
  labs(title = "BMI by smoke/gas exposure", x="BMI", y="n")
```

- Again, lots of people have normal BMI and exposure to gas.

```{r}
merged %>%
  ggplot() +
  geom_boxplot(mapping = aes(y=fev, fill=obesity_level))+
  labs(title = "FEV by BMI category", y="FEV")

merged %>%
    filter(!(smoke_gas_exposure %in% NA)) %>%
      ggplot() +
      geom_boxplot(mapping = aes(y=fev, fill=smoke_gas_exposure))+
      labs(title = "FEV by smoke/gas exposure", y="FEV")
```

- FEV again shows positive correlation with BMI but no clear correlation with gas/smoke exposure.

```{r}
leaflet(merged) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat = ~lat, lng = ~lon, color = "brown",opacity = 1,
             fillOpacity = 1, radius = ~(pm25_mass*200))
```

- PM2.5 mass is higher in greater LA area than in other Southern California communities.

```{r}
merged %>%
  ggplot(mapping = aes(x=pm25_mass, y=fev))+
  geom_point()+
  stat_smooth(method=lm)+
  labs(title = "PM2.5 mass v. FEV by town with regression lines")
```

- Since pm2.5 is only reported per town, it's hard to tell directly from this graph whether PM2.5 mass is associated with FEV.

```{r}
merged[, fev_avg_town := mean(fev, na.rm = TRUE), by = .(townname)]
merged %>%
  ggplot(mapping = aes(x=pm25_mass, y=fev_avg_town))+
  geom_point()+
  stat_smooth(method=lm)+
  labs(title = "PM2.5 mass v. FEV by town with regression lines")
```

- If we use the town average FEV to draw the graph instead, the PM2.5 mass shows negative association with FEV.