---
title: "Assignment 03 - Web Scraping and Text Mining"
output: github_document
---

```{r setup, message=FALSE}
library(httr)
library(stringr)
library(tidyverse)
library(data.table)
library(tidytext)
library(dplyr)
```


## APIs

```{r}
# Downloading the website 
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(
  website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

- I am able to find 564 papers.

```{r}
# Download details using query
query_ids <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db = "pubmed", term = "sars-cov-2 trial vaccine", retmax = 250))

# Extracting the content of the response of GET
ids <- httr::content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>") [[1]]

# Remove all the leading and trailing <Id> </Id>
ids <- stringr::str_remove_all(ids, "</?Id>")

# Get abstracts using ids
publications <- GET(
  url = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed", id = paste(ids, collapse = ","), retmax = 250, rettype = "abstract"))
```

```{r}
publications <- httr::content(publications)
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)

titles <- str_extract(pub_char_list,
                      "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")

journals <- str_extract(pub_char_list,
                       "<Title>(\\n|.)+</Title>")
journals <- str_remove_all(journals, "</?[[:alnum:]]+>")
journals <- str_replace_all(journals, "\\s+", " ")

dates <- str_extract_all(pub_char_list,
                             "<PubDate>(\\n|.)+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]]+>")
dates <- str_replace_all(dates, "\\s+", " ")

abstracts <- str_extract(pub_char_list,
                         "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+", " ")

database <- data.frame(
  PubMedID = ids, Title = titles, Journal = journals, Date = dates, Abstracts = abstracts)

knitr::kable(database)
```


## Text Mining

```{r, message=FALSE}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv",
  "pubmed.csv", method="libcurl", timeout = 60)

pubmed <- fread("pubmed.csv")
```

```{r}
pubmed %>%
  unnest_tokens(token, abstract) %>%
    count(token, sort = TRUE) %>%
      top_n(20)

pubmed %>%
  unnest_tokens(token, abstract) %>%
    anti_join(stop_words, by = c("token" = "word")) %>%
      filter(!(token %in% as.character(seq(0,100)))) %>%
        count(token, sort = TRUE) %>%
          top_n(20)

pubmed %>%
  filter(term == "covid") %>%
    unnest_tokens(token, abstract) %>%
      anti_join(stop_words, by = c("token" = "word")) %>%
        filter(!(token %in% as.character(seq(0,100)))) %>%
          count(token, sort = TRUE) %>%
            top_n(5)

pubmed %>%
  filter(term == "meningitis") %>%
    unnest_tokens(token, abstract) %>%
      anti_join(stop_words, by = c("token" = "word")) %>%
        filter(!(token %in% as.character(seq(0,100)))) %>%
          count(token, sort = TRUE) %>%
            top_n(5)

pubmed %>%
  filter(term == "prostate cancer") %>%
    unnest_tokens(token, abstract) %>%
      anti_join(stop_words, by = c("token" = "word")) %>%
        filter(!(token %in% as.character(seq(0,100)))) %>%
          count(token, sort = TRUE) %>%
            top_n(5)

pubmed %>%
  filter(term == "cystic fibrosis") %>%
    unnest_tokens(token, abstract) %>%
      anti_join(stop_words, by = c("token" = "word")) %>%
        filter(!(token %in% as.character(seq(0,100)))) %>%
          count(token, sort = TRUE) %>%
            top_n(5)

pubmed %>%
  filter(term == "preeclampsia") %>%
    unnest_tokens(token, abstract) %>%
      anti_join(stop_words, by = c("token" = "word")) %>%
        filter(!(token %in% as.character(seq(0,100)))) %>%
          count(token, sort = TRUE) %>%
            top_n(5)
```

- Removing stop words significantly change the tokens appear as the most frequent; only "covid" stays in top 10 after removing stop words. 

- After removing stop words, the 5 most common tokens for the search term:
    1) "covid" are "covid", "patients", "disease", "pandemic", "coronavirus" and "health" (6 because of ties);
    2) "meningitis" are "patients", "meningitis", "meningeal", "csf" and "clinical";
    3) "prostate cancer" are "cancer", "prostate", "patients", "treatment" and "disease";
    4) "cystic fibrosis" are "fibrosis", "cystic", "cf", "patients" and "disease";
    5) "preeclampsia" are "pre", "eclampsia", "preeclampsia", "women" and "pregnancy".
    
```{r}
pubmed %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
    count(ngram, sort = TRUE) %>%
      top_n(20) %>%
        ggplot(aes(x = n, y = fct_reorder(ngram, n))) + geom_col()
```

```{r}
tfidf <- pubmed %>%
  unnest_tokens(token, abstract) %>%
    count(token, term) %>%
      bind_tf_idf(token, term, n) %>%
        arrange(desc(tf_idf))

tfidf %>% 
  filter(term == "covid") %>%
    top_n(5)

tfidf %>% 
  filter(term == "meningitis") %>%
    top_n(5)

tfidf %>% 
  filter(term == "prostate cancer") %>%
    top_n(5)

tfidf %>% 
  filter(term == "cystic fibrosis") %>%
    top_n(5)

tfidf %>% 
  filter(term == "preeclampsia") %>%
    top_n(5)
```

- The 5 tokens with the highest TF-IDF value for the search term:
    1) "covid" are "covid", "pandemic", "coronavirus", "sars" and "cov";
    2) "meningitis" are "meningitis", "meningeal	", "pachymeningitis", "csf" and "meninges";
    3) "prostate cancer" are "prostate", "androgen", "psa", "prostatectomy" and "castration";
    4) "cystic fibrosis" are "cf", "fibrosis", "cystic", "cftr" and "sweat";
    5) "preeclampsia" are "eclampsia", "preeclampsia", "pregnancy", "maternal" and "gestational".
    
- The results are somewhat similar with the results in Q1, while become more in detail about the diseases.